---
title: 'Lab 9: Sentiment Analysis'
author: "Shuying Yu"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#Attach packages
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```


# Get The Hobbit

```{r, cache = TRUE}
hobbit_text <- pdf_text(here::here('data', 'the-hobbit.pdf'))

hobbit_p34 <- hobbit_text[34]
```


# Break down into pages and lines

```{r}
hobbit_lines <- data.frame(hobbit_text) %>% 
  mutate(page = 1:n()) %>%
  mutate(text_full = str_split(hobbit_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  
  #Get rid of extra white space
  mutate(text_full = str_trim(text_full)) 

```


# Do some tidying

Chapter 1 starts on line 138, so drop everything before it

```{r}
hobbit_chapts <- hobbit_lines %>% 
  
  #Remove row
  slice(-(1:137)) %>% 
  
  #Create chapter column
  mutate(chapter = ifelse(str_detect(text_full, "Chapter"), text_full, NA)) %>% 
  
  #Fill the NAs with everything before it (chapter number)
  fill(chapter, .direction = 'down') %>% 
  
  #Take chapter column, separate the chapter work and the number
  separate(col = chapter, into = c("ch", "no"), sep = " ") %>% 
  
  #Change roman numeral as a factor and write as a latin number
  mutate(chapter = as.numeric(as.roman(no)))
```


# Do some word counts by chapter

Tokenize text, break down text into tokens of choosing

We define token as individual words (others can sort by are sentences, *n*grams/bigrams, different ways of breaking them down)

```{r}
hobbit_words <- hobbit_chapts %>% 
  
  #Tokenize by words, drops punctuation
  unnest_tokens(word, text_full, token = "words") %>% 
  
  #Remove hobbit text
  select(-hobbit_text)




#Get total word count for each word by chapter
hobbit_wordcount <- hobbit_words %>% 
  count(chapter, word)

#The, and, etc are stop words
```


# Remove stop words

We can remove stop words

```{r}
#Tidyverse considers these as stop words, 1149 stop words
head(stop_words)


#Create one without stop words
hobbit_words_clean <- hobbit_words %>% 
  
  #Start with hobbit_words, remove anything in df that i want to join
  anti_join(stop_words, by = 'word')

```


Then let's try counting them again

```{r}
nonstop_counts <- hobbit_words_clean %>% 
  count(chapter, word)
```


## Find the top 5 words from each chapter

```{r}
#Top 5 words for each chapter
top_5_words <- nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5) %>%
  ungroup()
 
#Make some graphs: 
ggplot(data = top_5_words, aes(x = n, y = word)) +
  geom_col(fill = "blue") +
  
  #Scales = free means we see relative amount of each chapter
  #Each scale associated with each chapter
  facet_wrap(~chapter, scales = "free")
```

## Word cloud for Chapter 1

Built into `ggwordcloud` package

```{r}
ch1_top100 <- nonstop_counts %>% 
  filter(chapter == 1) %>% 
  
  #Arrange decending orders
  arrange(-n) %>% 
  
  #Just keep top 100
  slice(1:100)


#Create word cloud plot
ch1_cloud <- ggplot(data = ch1_top100, 
                    aes(label = word)) +
  
  #Define word cloud
  #Color and size changes based on n
  geom_text_wordcloud(aes(color = n, size = n), 
                      
                      #Change cloud based on shape
                      shape = "diamond") +
  
  #Scale sizes of words
  scale_size_area(max_size = 6) +
  
  #Make scale for color
  scale_color_gradientn(colors = c("darkgreen","blue","purple")) +
  
  #Change theme
  theme_minimal()
 


#Plot it
ch1_cloud
```

# Sentiment Analysis

Each word associated with emotional weight, positive or negative association.

The three general-purpose lexicons are

  -  AFINN from Finn Årup Nielsen, (positive and negative, weighted by strength of positive or negative association)
  -  bing from Bing Liu and collaborators, and (positive or negative, or 0)
  -  nrc from Saif Mohammad and Peter Turney (word emotion association lexicon, one of 8 emotions that the word is associated with)

All three of these lexicons are based on unigrams, i.e., single words. These lexicons contain many English words and the words are assigned



"afinn": Words ranked from -5 (very negative) to +5 (very positive)


```{r}
#Get sentiment data from tidyverse package
#Tibble of ~2500 words
#get_sentiments(lexicon = "afinn")
 
#Let's look at the pretty positive words:
afinn_pos <- get_sentiments("afinn") %>% 
  
  #Looks at 3,4,5 values
  filter(value %in% c(3,4,5))
```


For comparison, check out the bing lexicon: 

```{r}
#get_sentiments(lexicon = "bing")
```


Now nrc:
```{r}
#get_sentiments(lexicon = "nrc")
```



## Sentiment analysis with afinn: 

First, bind words in `hobbit_nonstop_words` to `afinn` lexicon:
```{r}
hobbit_afinn <- hobbit_words_clean %>% 
  
  #Keep instance of first df that match 2nd df
  #Keep rows that match between the two
  inner_join(get_sentiments("afinn"), by = 'word')
```



Let's find some counts (by sentiment ranking):


```{r}
#Howmany times a value shows up for each chapter
afinn_counts <- hobbit_afinn %>% 
  count(chapter, value)
 
#Plot them
ggplot(data = afinn_counts, aes(x = value, y = n)) +
  geom_col() +
  facet_wrap(~chapter)



#Find the mean afinn score by chapter: 
afinn_means <- hobbit_afinn %>% 
  group_by(chapter) %>% 
  summarize(mean_afinn = mean(value))
 

#Plot it
ggplot(data = afinn_means, 
       aes(x = fct_rev(factor(chapter)),
           y = mean_afinn)) +
  
  #Chapter 1 on top, so force as factor to be 1-20
           # y = fct_rev(as.factor(chapter)))) +
  geom_col() +
  
  coord_flip()
```

## Now with NRC lexicon

Recall, this assigns words to sentiment bins. Let's bind our hobbit data to the NRC lexicon: 

```{r}
hobbit_nrc <- hobbit_words_clean %>% 
  inner_join(get_sentiments("nrc"))
```


Let's find the count of words by chapter and sentiment bin: 

```{r}

#count chapter and sentiment
hobbit_nrc_counts <- hobbit_nrc %>% 
  count(chapter, sentiment)
 

#Plot it

#aes(x = sentiment, y = n)
ggplot(data = hobbit_nrc_counts, aes(x = chapter, y = n)) +
  geom_col() +
  #facet_wrap(~chapter) +
  
  facet_wrap(~ sentiment)
  coord_flip()
```
















